Scheduling and Amortization for MPC

### 2021-11-29 paper_SIMD.tex
Fixed bugs in paper.
Added removal of edges.

\ana{Have to double check and make sure we capture all the cases.}
The dependence graph has the following def-use edges:

@@

>>>>

\ana{We need to double check that the following captures all cases of def-use edges.}
The dependence graph has the following def-use edges:

@@

\subsubsection{Removal of Infeasible Edges}

Array writes limit vectorization as they introduce sometimes infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}

\begin{algorithmic}
\STATE for i in range(N):
\STATE \hspace{0.25cm} A[i] = B[i] + 10;
\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
\end{algorithmic}

In Cytron's SSA this code (roughly) translates into

\begin{algorithmic}
\STATE for i in range(N):
\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$ 
\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$ 
\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$ 
\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$ 
\STATE 6. $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10;
\STATE 7. $\texttt{B}_1$ = $\texttt{B}_0$ 
\STATE 8. $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
\STATE 9. $\texttt{C}_1$ = $\texttt{C}_0$ 
\STATE 10. $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
\STATE 11. $\texttt{D}_1$ = $\texttt{D}_0$ 
\STATE 12. $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
\end{algorithmic}

There is a cycle around $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$ that includes statement $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10;
and that statement won't be vectorized even though in fact there is no loop-carried dependency from the write of $\texttt{B}_1$[i] at 8 to the 
read of $ ... = \texttt{B}_0$[i] at 6.

The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.

\begin{algorithmic}
\FOR {each array $\texttt{A}$ written in loop $j$} 
\STATE \COMMENT { including enclosed loops in $j$ }
\STATE dep = False
\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
\IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t. $0\! \le\! \underline{i}\! <\! I,\: 0\! \le\! \underline{j}, \underline{j}' \! <\! J,\: 0\! \le\! \underline{k}, \underline{k}' \! <\! K,\: \underline{j}<\underline{j}'$, and $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$}
\STATE dep = True
\ENDIF
\ENDFOR
\IF {dep == False}
\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
\ENDIF
\ENDFOR
\end{algorithmic}

Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$ 
references the \emph{same} array element as an use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}$, 
we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy 
$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential 
loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. \ana{Double check this!!!}

Consider the earlier example. There is a single loop, $i$. Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$ (due to the def-use pairs of \texttt{A} 6-8 and 6-10).
Therefore, we remove the back edge from 6 to 1. Analogously, we remove the back edges from 8 to 2 and 10 to 3. However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 12 to 4 remains (def-use pairs for \texttt{D}). As a result of removing these spurious edges, Basic Vectorization will find that statement 6 is vectorizable. Statements 8, 10 and 12 will correctly appear in the FOR loop.

\subsubsection{Array MUX refinement}

Next, the algorithm refines array MUX statement. 
MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index 
of A can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single 
index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$. \ana{Clarify why this helps.}

\begin{algorithmic}
\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
\IF {$i_1 \cong i_2$} \STATE \COMMENT{ Also $i_1 \cong i_2$ when one is null. WLG assume $i_2$ is null } 
\STATE replace $\mathit{stmt}$ with \\
\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
%\ELSIF {$i_1 \ncong i_2$}  
%\STATE replace $\mathit{stmt}$ with \\
%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
\ELSE \STATE \COMMENT{ Also, not $\cong$ when one is $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$}
\STATE $\mathit{stmt}$ stays as is
\ENDIF
\ENDFOR
\end{algorithmic}
